Security Properties for Stack Safety

### 2021-05-14 short/main.tex
Discussion from today and some additional detail

We propose such a characterization,
using the technical framework of language-based security.
Stack safety protects a caller from its callee, guaranteeing
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control. This formulation not only captures the intuition
that the callee cannot directly access the caller's state, but gives a novel
way of looking at control-flow attacks such as return-oriented-programming (ROP).
In an ROP attack, the callee pretends to return to its caller, thus accessing the
caller's privileges, but does so in such a way that it is still in control of
execution. Our model allows the caller to dictate the terms of a valid return --
typically, that the stack pointer is restored and execution proceeds from the
return address. An ROP attack that ``returns'' and reads the caller's state is
therefore treated precisely as if the callee had never returned, but attempted
to read the caller's state directly. It is a confidentiality violation. Likewise
an attack that would corrupt the caller's state is an integrity violation.

We formalize integrity and confidentiality as trace properties, in two different
variants: {\em stepwise} variants, in which a caller's data is {\em never}
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional, and
any reasonable protection mechanism ought to enforce them,
even if it does not prevent every single dangerous read or write.
Confidentiality is especially interesting, as it is based on the traditional
notion of noninterference. But where noninterference is normally presented as an
end-to-end hyperproperty, stack confidentiality is noninterference applied over
multiple nested subtraces -- one for every call.

To demonstrate the utility of our properties, we use them
to validate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that these micro-policies correctly abort programs that
would violate stack safety.

Our testing supports the stack-safety claim of Roessler and Dehon's {\em Depth
  Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing. On the other hand, we find that their \emph{Lazy
  Tagging and Clearing} policy
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that does enforce
confidentiality (and test that it does), albeit at some performance cost.

Finally, we demonstrate our model's flexibility by extending it with the passing
of variables on the stack and to a simple coroutine model.

>>>>

We propose such a characterization,
using the technical framework of language-based security.
Stack safety protects a caller from its callee, guaranteeing
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control.

We formalize integrity and confidentiality as trace properties, in two different
variants: {\em stepwise} variants, in which a caller's data is {\em never}
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional, and
any reasonable protection mechanism ought to enforce them,
even if it does not prevent every single dangerous read or write.

%SNA: will this audience know "hyperproperty?"
Confidentiality is especially interesting, as it is based on the traditional
notion of noninterference. But where noninterference is normally presented as an
end-to-end hyperproperty, stack confidentiality is noninterference applied over
multiple nested subtraces -- each individual call's behavior must be invariant
regardless of the state of the stack at its entry.

This formulation not only captures the intuition
that the callee cannot directly access the caller's state, but gives a novel
way of looking at control-flow attacks. We are not interested in attacks that
merely aim to execute arbitrary code -- in fact, our attacker model assumes
that the (attacking) callee may already execute arbitrary code.
%
By extension, we do not place any particular restrictions on control flow.
Perhaps suprisingly, we can still meaningfully distinguish the caller from
its callee, and talk about protecting the caller's data. After a call, control
is considered to belong to the callee until the trace reaches a valid
``return target,'' typically a state with the program counter at the next instruction
and the stack pointer restored. If the caller's data are accessed before then,
it is a violation of integrity or confidentiality.

These properties can optionally be extended with a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}, the system-wide control-flow property that callees always return
to their immediate caller, if they return at all. This property is largely orthogonal to our
data-protection properties.

To demonstrate the utility of our properties, we use them
to validate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that these micro-policies correctly abort programs that
would violate stack safety. Furthermore, we check incorrect
enforcement variants (both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself). The testing framework
is able to generate counterexamples that violate our properties but are not halted by
incorrect enforcement mechanisms, increasing our confidence when it finds
no counterexamples under the real enforcement mechanisms.

Our testing supports the stack-safety claim of Roessler and Dehon's {\em Depth
  Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the depth of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing.
On the other hand, we find that their \emph{Lazy Tagging and Clearing} policy
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that should enforce observational
integrity and confidentiality, albeit at some performance cost. The next step
is to expand testing to these properties and validate the variant; efficiently
testing observational properties presents unique challenges because a property
violation can be arbitrarily far removed from the initial ``risky'' read or write.

Finally, we demonstrate our model's flexibility by extending it to different settings.
First, we allow variables to be passed on the stack, allowing some of the state of
the caller to be shared with the callee (and potentially with a further
callee, and so on.) Then we expand the number of stacks as we move to a simple
coroutine model, albiet one with statically bounded regions for each stack.

In addition to testing lazy tag-based policies against our observational properties,
we are adapting the uninitialized capability model of \citet{Georges+21} to work with
our testing framework for validation against our properties. Of the several Cheri-based
stack protection schemes, its treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. There is no equivalent dynamic requirement in a tag-based system because
the policy can dynamically enforce cooperation.
